Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-01-14 jss/EpiCompare-paper/EpiCompare-paper.Rmd
Section 1 and 2 changes and references

The recent (and on-going) COVID-19 global pandemic has galvanized public interest in understanding more about infectious disease modeling and has highlighted the usefulness of research in the area of infectious disease epidemiology. Infectious diseases inflict enormous burdens on the world: millions of lives lost and trillions of dollars spent yearly. Infectious disease models typically attempt to do one or more of the following: 1) predict the spread of current and future epidemics \citep[e.g. flu prediction][]{Biggerstaff2016}, 2) analyze past and current epidemics to increase scientific knowledge \citep[e.g. historical measle outbreaks][]{Neal2004}, and 3) forecast or project epidemic scenarios under pre-specified parameters \citep[e.g. ...][]{}^[\@Shannon: CITE]. At the same time, descriptive statistics and visualizations (from universities like John Hopkins \citep{}^[\@Shannon: CITE], many branches and levels of government, and new organizations \citep{}^[\@Shannon: CITE]) are an important first step of the process.

With the many visualization and exploratory tools, models and modeling paradigms, and reviews and comparisons in the literature and through MIDAS (Models of Infectious Disease Agent Study) Control Center comparison guides \citep{}^[\@Shannon: CITE], this field has a lot of devices to aid an individual practitioner decide the correct approach. At the same time, analysis packages often only address a portion of the analysis pipeline. Modeling tools often don't provide easy ways to compare and assess their models on new data. Moreover, exploring and modeling epidemics require transforming and \textit{tidying} data in different ways. To fill these gaps, we present our \proglang{R} package \pkg{EpiCompare}. Our package's primary focus is to aid and advance research in the area of comparison and assessment of epidemic \& epidemiological models. In Figure \ref{fig:pipeline}, we illustrate the data analysis pipeline of infectious diseases  as 1) data pre-processing, 2) exploratory data analysis (EDA), 3) modeling and simulating, 4) post-processing, and 5) comparison and assessment; where each previous part of the pipeline influences the next. \pkg{EpiCompare} provides tools to aids practitioners in all areas of this pipeline.

@@

This package also emphasizes the value of approach epidemics in a \textit{time-invariant} way, not constrained to things like a specific time scale, or initial date of the first infection. Although epidemics are by definition a process that evolves over time, epidemics, including the recent COVID-19, often need to be compared in a time-invariant way to understand the processes at play. Additionally, many tools to examine the quantity of the population in each state along the infection process (for example: quantity of suspectible vs infected vs recovered individuals) don't always as intelligently capture the natural connections between the proportion of individuals in these states. Tools in \pkg{EpiCompare} attempt to give the user the ability to extend their toolkit to evaluate epidemics to also include time-invariant approaches. The goal of \pkg{EpiCompare} is not to supplant existing infectious disease modeling tools and software but, rather, is a concerted effort to create standard and fair comparisons among models developed for disease outbreaks and outbreak data. \shannon{flesh out time invariance introduction to better explain to non-experts what we mean.  perhaps mention R0}

@@

SHANNON: I THINK JUST INCLUDING LOTS OF LITERATURE IN INTRO + ETC. INSTEAD OF HAVING A LITERATURE REVIEW IS BETTER.

@@

Epidemics can be difficult to compare to one another due to differences in diseases, locations, or population behaviors, or times.  We focus on comparisons between epidemics adjusting for time, both the relative unit of rate of infection (e.g. days, months, years) and the relative beginning and end of an epidemic.  By adjusting for the unit of infection rate, we can focus on the ``lifetime'' of an epidemic, a view that is concerned more with the total number of lives affected than due to a time constraint.  By adjusting for the relative beginning/end of an infection we can compare, for example the first part of infection from one city to another where this is some time gap in spread.

One way to adjust for both facets of time may seem obvious -- scale the time between zero and one, based on the initial time of infection $t_0$ and the final time $t_F$.  Upon inspection, this method of adjustment requires clear definition of what a beginning and end of an epidemic constitute which requires making one or more choices.  We propose the use of time-invariant analysis as it appears in \pkg{EpiCompare}, which avoids the need of making such a choice.

## What $R_0$ has to do with time-invariant analysis {short-title="r0" #r0}

 By definition, $R_0$ is the number of expected secondary infections when a primary infection is introduced to a susceptible population.  $R_0$ is also, maybe, the most famous \textit{time-invariant} numerical summary of an epidemic, which allows epidemics to be compared to one another in both different time and geographic scales. For example, $R_0$ for Covid-19 is estimated to be between 2-3, seasonal influenza between 1.2-2, and modern measles outbreaks between 5-6.

Estimator(s) for $R_0$ are dependent on the epidemic modeling framework, which consists of which states an individual can occupy (e.g. susceptible, infectious, recovered) and a description of how individuals move from one state to the next over time.  A common epidemic modeling framework is the SIR model, originally introduced by @kermack1927.  Transitions from one state to the next are defined by a series of ordinary differential equations, where $N$ is the (fixed) total number individuals, $\beta$ is the rate of infection, and $\gamma$ is rate of recovery,
\begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align}
From this, $\hat{R}_0 = \frac{\hat{\beta}}{\hat{\gamma}}$, the ratio of the estimated infection rate compared to the estimated recovery rate.

@@

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent result from \cite{Harko2014} and is shown in detail in Appendix \ref{app:proof}.  The consequence of Theorem \ref{thm:sir-scale} is that for two SIR models that have the same initial percent of individuals in each state and $R_0$ then for every point on the epidemic path of the first SIR model is also a point on the epidemic path of the second SIR model. Taking the sample simulations from Fig. \ref{fig:different-scales-standard}, Fig. \ref{fig:different-scales-tern} presents these two models in a ternary plot.  This means, that at a glance, we can tell if two epidemics have different values of $R_0$.

@@

In this section, we highlight a number of the functionalities available in \proglang{EpiCompare}.  These functionalities include data cleaning, visualization, simulation, and comparison, in accordance with the data analysis pipeline REF\ref{}. We show a full data analysis from beginning to end that can be accomplished in a streamlined and standardized manner.

>>>>

The recent (and on-going) COVID-19 global pandemic has galvanized public interest in understanding more about infectious disease modeling and has highlighted the usefulness of research in the area of infectious disease epidemiology. Infectious diseases inflict enormous burdens on the world: millions of lives lost and trillions of dollars spent yearly. Infectious disease models typically attempt to do one or more of the following: 1) predict the spread of current and future epidemics \citep[e.g. flu prediction][]{Biggerstaff2016}, 2) analyze past and current epidemics to increase scientific knowledge \citep[e.g. historical measle outbreaks][]{Neal2004}, and 3) forecast or project epidemic scenarios under pre-specified parameters \citep[e.g.][]{ferguson2020}. At the same time, descriptive statistics and visualizations from universities, many branches and levels of government, and news organizations  are an important first step of the process \citep{dong2020,cdc-covid-tracker2021,wp-covid-tracker2021} .

With the many visualization and exploratory tools, models and modeling paradigms, and reviews and comparisons in the literature and through the MIDAS (Models of Infectious Disease Agent Study) network \citep{midasnetwork2021}, this field has a lot of devices to aid an individual practitioner decide the correct approach.  For example,\proglang{R} packages such as \pkg{surveillance}, \pkg{EpiModel}, and \pkg{pomp} have all made significant steps in standardizing the flow of the data analysis pipeline for epidemic modeling through digitizing data sets, making accessible statistical models, and providing a plethora of educational material for both coding novices and experts alike \citep{surveillance2017,Jenness2018,King2016}.

At the same time, analysis packages often only address a specific portion of the analysis pipeline, for instance focusing on certain types of models. Modeling tools, which usually require learning package-specific syntax, often don't provide easy ways to compare and assess their models on new data. Moreover, exploring and modeling epidemics require transforming and \textit{tidying} data in different ways. To fill these gaps, we present our \proglang{R} package \pkg{EpiCompare}. Our package's primary focus is to aid and advance research in the area of comparison and assessment of epidemic and epidemiological models. In Figure \ref{fig:pipeline}, we illustrate the data analysis pipeline of infectious diseases  as 1) data pre-processing, 2) exploratory data analysis (EDA), 3) modeling and simulating, 4) post-processing, and 5) comparison and assessment; where each previous part of the pipeline influences the next. \pkg{EpiCompare} provides tools to aids practitioners in all areas of this pipeline.

@@

\pkg{EpiCompare} also emphasizes the value of analyzing epidemics in a \textit{time-invariant} way. Epidemics, despite by definition being a process that evolves over time, often need to be compared in a way not constrained to initial times or time scales to understand the processes at play. Additionally, many tools designed to examine the quantity of the population in each epidemic state (for example: quantity of Susceptible vs Infectious vs Recovered individuals) don't always as intelligently capture the natural connections between the proportion of individuals in these states. \shannon{I don't quite understand this.  Is this because of underreporting?  If so I'm not sure how EpiCompare helps us here.} Tools in \pkg{EpiCompare} give the user the ability to extend their toolkit to evaluate epidemics within a time-invariant lens. The goal of \pkg{EpiCompare} is not to supplant existing infectious disease modeling tools and software but, rather, is a concerted effort to create standard and fair comparisons among models developed for disease outbreaks and outbreak data.

@@

@@

Epidemics can be difficult to compare to one another due to differences in diseases, locations, or population behaviors, or times.  In \pkg{EpiCompare}, we emphasize comparisons between epidemics adjusting for the last component, time.  Time-invariant analysis is beneficial because by adjusting for the unit of infection rate, we can focus on the "lifetime" of an epidemic, a view that is concerned more with the number of lives affected than as opposed to any specific time constraint.    Time-invariant analysis is also beneficial when there are gaps in time between occurrences of outbreaks of a similar nature in different geographic regions.  Finally, time-invariant analysis is beneficial because many studies talk about a period of ``exponential growth'' of the number of outbreaks \citep{chowell2007,wallinga2007generation,forsberg2008}.

Time-invariant analysis, as it appears in \pkg{EpiCompare} solves the above issues by observing functionals which attempt to capture the general shape of the epidemic with respect to the proportion of the population in each epidemic state.  This allows us to compare scenarios as different as, for instance as a decades-long outbreak HIV in the US compared to a 10 day outbreak of norovirus on a cruise ship.  Moreover, this tool avoids the need for choosing a 'beginning' $t_0$ or 'end' $t_F$ of an epidemic, choices \cite{gallagher2020} show that can heavily influence, for example,  estimates of peak infection height or the reproduction number $R_0$.

## $R_0$  and time-invariant analysis {short-title="r0" #r0}

 By definition, $R_0$ is the number of expected secondary infections when a primary infection is introduced to a susceptible population.  $R_0$ is also, maybe, the most famous \textit{time-invariant} numerical summary of an epidemic, which allows epidemics to be compared to one another in both different time and geographic scales. For example, $R_0$ for Covid-19 is estimated to be between 2-3, seasonal influenza between 1.2-2, and modern measles outbreaks between 5-6 \citep{midas2020,biggerstaff2014,namee2018}.

Estimator(s) for $R_0$ are dependent on the epidemic modeling framework, which consists of which states an individual can occupy (e.g. susceptible, infectious, recovered) and a description of how individuals move from one state to the next over time (see \cite{hethcote1994}).  A common epidemic modeling framework is the SIR model, originally introduced by @kermack1927.  Transitions from one state to the next are defined by a series of ordinary differential equations, where $N$ is the (fixed) total number individuals, $\beta$ is the rate of infection, and $\gamma$ is rate of recovery,
\begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align}
From this, $\hat{R}_0 = \frac{\hat{\beta}}{\hat{\gamma}}$, the ratio of the estimated infection rate compared to the estimated recovery rate.

@@

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent result from \cite{Harko2014} and is shown in detail in Appendix \ref{app:proof}.  The consequence of Theorem \ref{thm:sir-scale} is that for two SIR models that have the same initial percent of individuals in each state and $R_0$ then for every point on the epidemic path of the first SIR model is also a point on the epidemic path of the second SIR model. Taking the sample simulations from Fig. \ref{fig:different-scales-standard}, Fig. \ref{fig:different-scales-tern} presents these two models in a ternary plot.  This means with our time-invariant ternary plot, that at a glance, we can tell if two epidemics have different values of $R_0$.

@@

In this section, we highlight a number of the functionalities available in \kg{EpiCompare}.  These functionalities include data cleaning, visualization, simulation, and comparison, in accordance with the data analysis pipeline REF\ref{}. We show a full data analysis from beginning to end that can be accomplished in a streamlined and standardized manner.

### 2021-01-14 jss/EpiCompare-paper/EpiCompare-paper.tex
Section 1 and 2 changes and references

The recent (and on-going) COVID-19 global pandemic has galvanized public
interest in understanding more about infectious disease modeling and has
highlighted the usefulness of research in the area of infectious disease
epidemiology. Infectious diseases inflict enormous burdens on the world:
millions of lives lost and trillions of dollars spent yearly. Infectious
disease models typically attempt to do one or more of the following: 1)
predict the spread of current and future epidemics
\citep[e.g. flu prediction][]{Biggerstaff2016}, 2) analyze past and
current epidemics to increase scientific knowledge
\citep[e.g. historical measle outbreaks][]{Neal2004}, and 3) forecast or
project epidemic scenarios under pre-specified parameters
\citep[e.g. ...][]{}\footnote{@Shannon: CITE}. At the same time,
descriptive statistics and visualizations (from universities like John
Hopkins \citep{}\footnote{@Shannon: CITE}, many branches and levels of
government, and new organizations \citep{}\footnote{@Shannon: CITE}) are
an important first step of the process.

With the many visualization and exploratory tools, models and modeling
paradigms, and reviews and comparisons in the literature and through
MIDAS (Models of Infectious Disease Agent Study) Control Center
comparison guides \citep{}\footnote{@Shannon: CITE}, this field has a
lot of devices to aid an individual practitioner decide the correct
approach. At the same time, analysis packages often only address a
portion of the analysis pipeline. Modeling tools often don't provide
easy ways to compare and assess their models on new data. Moreover,
exploring and modeling epidemics require transforming and
\textit{tidying} data in different ways. To fill these gaps, we present
our \proglang{R} package \pkg{EpiCompare}. Our package's primary focus
is to aid and advance research in the area of comparison and assessment
of epidemic \& epidemiological models. In Figure \ref{fig:pipeline}, we
illustrate the data analysis pipeline of infectious diseases as 1) data
pre-processing, 2) exploratory data analysis (EDA), 3) modeling and
simulating, 4) post-processing, and 5) comparison and assessment; where
each previous part of the pipeline influences the next. \pkg{EpiCompare}
provides tools to aids practitioners in all areas of this pipeline.

@@

This package also emphasizes the value of approach epidemics in a
\textit{time-invariant} way, not constrained to things like a specific
time scale, or initial date of the first infection. Although epidemics
are by definition a process that evolves over time, epidemics, including
the recent COVID-19, often need to be compared in a time-invariant way
to understand the processes at play. Additionally, many tools to examine
the quantity of the population in each state along the infection process
(for example: quantity of suspectible vs infected vs recovered
individuals) don't always as intelligently capture the natural
connections between the proportion of individuals in these states. Tools
in \pkg{EpiCompare} attempt to give the user the ability to extend their
toolkit to evaluate epidemics to also include time-invariant approaches.
The goal of \pkg{EpiCompare} is not to supplant existing infectious
disease modeling tools and software but, rather, is a concerted effort
to create standard and fair comparisons among models developed for
disease outbreaks and outbreak data.
\textcolor{orange}{flesh out time invariance introduction to better explain to non-experts what we mean.  perhaps mention R0}

@@

Epidemics can be difficult to compare to one another due to differences
in diseases, locations, or population behaviors, or times. We focus on
comparisons between epidemics adjusting for time, both the relative unit
of rate of infection (e.g.~days, months, years) and the relative
beginning and end of an epidemic. By adjusting for the unit of infection
rate, we can focus on the ``lifetime'\,' of an epidemic, a view that is
concerned more with the total number of lives affected than due to a
time constraint. By adjusting for the relative beginning/end of an
infection we can compare, for example the first part of infection from
one city to another where this is some time gap in spread.

One way to adjust for both facets of time may seem obvious -- scale the
time between zero and one, based on the initial time of infection
\(t_0\) and the final time \(t_F\). Upon inspection, this method of
adjustment requires clear definition of what a beginning and end of an
epidemic constitute which requires making one or more choices. We
propose the use of time-invariant analysis as it appears in
\pkg{EpiCompare}, which avoids the need of making such a choice.

\subsection[r0]{What \(R_0\) has to do with time-invariant
analysis}\label{r0}

By definition, \(R_0\) is the number of expected secondary infections
when a primary infection is introduced to a susceptible population.
\(R_0\) is also, maybe, the most famous \textit{time-invariant}
numerical summary of an epidemic, which allows epidemics to be compared
to one another in both different time and geographic scales. For
example, \(R_0\) for Covid-19 is estimated to be between 2-3, seasonal
influenza between 1.2-2, and modern measles outbreaks between 5-6.

Estimator(s) for \(R_0\) are dependent on the epidemic modeling
framework, which consists of which states an individual can occupy
(e.g.~susceptible, infectious, recovered) and a description of how
individuals move from one state to the next over time. A common epidemic
modeling framework is the SIR model, originally introduced by
\citet{kermack1927}. Transitions from one state to the next are defined
by a series of ordinary differential equations, where \(N\) is the
(fixed) total number individuals, \(\beta\) is the rate of infection,
and \(\gamma\) is rate of recovery, \begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align} From this,
\(\hat{R}_0 = \frac{\hat{\beta}}{\hat{\gamma}}\), the ratio of the
estimated infection rate compared to the estimated recovery rate.

@@

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent
result from \cite{Harko2014} and is shown in detail in Appendix
\ref{app:proof}. The consequence of Theorem \ref{thm:sir-scale} is that
for two SIR models that have the same initial percent of individuals in
each state and \(R_0\) then for every point on the epidemic path of the
first SIR model is also a point on the epidemic path of the second SIR
model. Taking the sample simulations from Fig.
\ref{fig:different-scales-standard}, Fig.
\ref{fig:different-scales-tern} presents these two models in a ternary
plot. This means, that at a glance, we can tell if two epidemics have
different values of \(R_0\).

>>>>

The recent (and on-going) COVID-19 global pandemic has galvanized public
interest in understanding more about infectious disease modeling and has
highlighted the usefulness of research in the area of infectious disease
epidemiology. Infectious diseases inflict enormous burdens on the world:
millions of lives lost and trillions of dollars spent yearly. Infectious
disease models typically attempt to do one or more of the following: 1)
predict the spread of current and future epidemics
\citep[e.g. flu prediction][]{Biggerstaff2016}, 2) analyze past and
current epidemics to increase scientific knowledge
\citep[e.g. historical measle outbreaks][]{Neal2004}, and 3) forecast or
project epidemic scenarios under pre-specified parameters
\citep[e.g.][]{ferguson2020}. At the same time, descriptive statistics
and visualizations from universities, many branches and levels of
government, and news organizations are an important first step of the
process \citep{dong2020,cdc-covid-tracker2021,wp-covid-tracker2021} .

With the many visualization and exploratory tools, models and modeling
paradigms, and reviews and comparisons in the literature and through the
MIDAS (Models of Infectious Disease Agent Study) network
\citep{midasnetwork2021}, this field has a lot of devices to aid an
individual practitioner decide the correct approach. For
example,\proglang{R} packages such as \pkg{surveillance},
\pkg{EpiModel}, and \pkg{pomp} have all made significant steps in
standardizing the flow of the data analysis pipeline for epidemic
modeling through digitizing data sets, making accessible statistical
models, and providing a plethora of educational material for both coding
novices and experts alike \citep{surveillance2017,Jenness2018,King2016}.

At the same time, analysis packages often only address a specific
portion of the analysis pipeline, for instance focusing on certain types
of models. Modeling tools, which usually require learning
package-specific syntax, often don't provide easy ways to compare and
assess their models on new data. Moreover, exploring and modeling
epidemics require transforming and \textit{tidying} data in different
ways. To fill these gaps, we present our \proglang{R} package
\pkg{EpiCompare}. Our package's primary focus is to aid and advance
research in the area of comparison and assessment of epidemic and
epidemiological models. In Figure \ref{fig:pipeline}, we illustrate the
data analysis pipeline of infectious diseases as 1) data pre-processing,
2) exploratory data analysis (EDA), 3) modeling and simulating, 4)
post-processing, and 5) comparison and assessment; where each previous
part of the pipeline influences the next. \pkg{EpiCompare} provides
tools to aids practitioners in all areas of this pipeline.

@@

\pkg{EpiCompare} also emphasizes the value of analyzing epidemics in a
\textit{time-invariant} way. Epidemics, despite by definition being a
process that evolves over time, often need to be compared in a way not
constrained to initial times or time scales to understand the processes
at play. Additionally, many tools designed to examine the quantity of
the population in each epidemic state (for example: quantity of
Susceptible vs Infectious vs Recovered individuals) don't always as
intelligently capture the natural connections between the proportion of
individuals in these states.
\textcolor{orange}{I don't quite understand this.  Is this because of underreporting?  If so I'm not sure how EpiCompare helps us here.}
Tools in \pkg{EpiCompare} give the user the ability to extend their
toolkit to evaluate epidemics within a time-invariant lens. The goal of
\pkg{EpiCompare} is not to supplant existing infectious disease modeling
tools and software but, rather, is a concerted effort to create standard
and fair comparisons among models developed for disease outbreaks and
outbreak data.

@@

Epidemics can be difficult to compare to one another due to differences
in diseases, locations, or population behaviors, or times. In
\pkg{EpiCompare}, we emphasize comparisons between epidemics adjusting
for the last component, time. Time-invariant analysis is beneficial
because by adjusting for the unit of infection rate, we can focus on the
``lifetime'' of an epidemic, a view that is concerned more with the
number of lives affected than as opposed to any specific time
constraint. Time-invariant analysis is also beneficial when there are
gaps in time between occurrences of outbreaks of a similar nature in
different geographic regions. Finally, time-invariant analysis is
beneficial because many studies talk about a period of ``exponential
growth'\,' of the number of outbreaks
\citep{chowell2007,wallinga2007generation,forsberg2008}.

Time-invariant analysis, as it appears in \pkg{EpiCompare} solves the
above issues by observing functionals which attempt to capture the
general shape of the epidemic with respect to the proportion of the
population in each epidemic state. This allows us to compare scenarios
as different as, for instance as a decades-long outbreak HIV in the US
compared to a 10 day outbreak of norovirus on a cruise ship. Moreover,
this tool avoids the need for choosing a `beginning' \(t_0\) or `end'
\(t_F\) of an epidemic, choices \cite{gallagher2020} show that can
heavily influence, for example, estimates of peak infection height or
the reproduction number \(R_0\).

\subsection[r0]{\(R_0\) and time-invariant analysis}\label{r0}

By definition, \(R_0\) is the number of expected secondary infections
when a primary infection is introduced to a susceptible population.
\(R_0\) is also, maybe, the most famous \textit{time-invariant}
numerical summary of an epidemic, which allows epidemics to be compared
to one another in both different time and geographic scales. For
example, \(R_0\) for Covid-19 is estimated to be between 2-3, seasonal
influenza between 1.2-2, and modern measles outbreaks between 5-6
\citep{midas2020,biggerstaff2014,namee2018}.

Estimator(s) for \(R_0\) are dependent on the epidemic modeling
framework, which consists of which states an individual can occupy
(e.g.~susceptible, infectious, recovered) and a description of how
individuals move from one state to the next over time (see
\cite{hethcote1994}). A common epidemic modeling framework is the SIR
model, originally introduced by \citet{kermack1927}. Transitions from
one state to the next are defined by a series of ordinary differential
equations, where \(N\) is the (fixed) total number individuals,
\(\beta\) is the rate of infection, and \(\gamma\) is rate of recovery,
\begin{align}\label{eq:sir-ode}
      S^\prime(t) &= -\frac{\beta S(t)I(t)}{N} \\
      I^\prime(t) &= \frac{\beta S(t)I(t)}{N} - \gamma I(t) \nonumber\\
      R^\prime(t) &= \gamma I(t) \nonumber.
  \end{align} From this,
\(\hat{R}_0 = \frac{\hat{\beta}}{\hat{\gamma}}\), the ratio of the
estimated infection rate compared to the estimated recovery rate.

@@

The proof of Theorem \ref{thm:sir-scale} relies on a fairly recent
result from \cite{Harko2014} and is shown in detail in Appendix
\ref{app:proof}. The consequence of Theorem \ref{thm:sir-scale} is that
for two SIR models that have the same initial percent of individuals in
each state and \(R_0\) then for every point on the epidemic path of the
first SIR model is also a point on the epidemic path of the second SIR
model. Taking the sample simulations from Fig.
\ref{fig:different-scales-standard}, Fig.
\ref{fig:different-scales-tern} presents these two models in a ternary
plot. This means with our time-invariant ternary plot, that at a glance,
we can tell if two epidemics have different values of \(R_0\).
